### 2024-02-18 1-layer.tex
Updated 2-layer after TV comments

\begin{center}
\begin{minipage}[b]{0.9\linewidth}
\small
\textbf{Foreword\,}
This chapter is an \emph{in extenso} reproduction of \cite{Bahl2023a} and has been submitted to Journal of Biophotonics for publication. 
\newline
AB identified the three major analytical tissue models for comparison, designed and conducted the simulations and synthetic experiments and measurements to construct the datasets for evaluation, and conducted the evaluation of the models under the supervision of JS, MB, and TV. AB worked collaboratively with SS and YX to identify and validate the appropriate IAD method to use in this work. 
\end{minipage}
\end{center}

@@

% \maketitle 
% 
\section{Introduction}\label{sec:intro}
As mentioned in Chapter \ref{chap:intro}, $StO_2$ 
%Tissue oxygen saturation ($StO_2$) 
is a key metric that could be used in surgery to determine tissue viability, tumour localisation, and vessel flow \cite{Takami2017, Hughes2019, Richardson2023}.
There are, however, no currently established, spatially-resolved, intra-operative methods of quantitatively determining $StO_2$ and so perfusion is often displayed (e.g. using DSA, MVD, or ICG) or global $StO_2$ is discussed from JVOS or cerebral oximetry.
% In many clinical applications indocyanine green fluorescence angiography (ICG) is used to display perfusion as 
% %oxygenated
% blood flow through vessels can be visualised by the fluorescence of the indocyanine dye given to the patient \cite{Hackethal2018}.
% Whilst this method has shown clinical benefits \cite{Renna2023, Teng2021}, it cannot quantify the $StO_2$ of the tissues being visualised leading to subjective interpretation.
There has been increasing promise that diffuse reflectance spectroscopic techniques could be used to obtain $StO_2$ intra-operatively. Hyperspectral Imaging (HSI) is one such technique that obtains a diffuse reflectance spectrum for each pixel of an image \cite{Kulcke2018, Taylor-Williams2022}, enabling intra-operative, non-contact, spatially-resolved $StO_2$ extraction without the need for a contrast agent.

There have been many proposed methods to extract $StO_2$ parameters from tissue diffuse reflectance spectra ranging in complexity, as mentioned in Chapter \ref{chap:intro}.
%from the simplest ratiometric two-wavelength methods to more sophisticated analytical models \cite{MacKenzie2018} or deep-learning based approaches \cite{Ayala2023}. Using two or three wavelength models can provide good results but require data to be captured at these precise wavelengths placing large constraints on the measurement devices used and significant model assumptions. 
Analytical models can be applied to ranges of wavelengths for which they are developed and so these are favoured in this work. 
%\revref{2}{6}\textcolor{red}{
Here we consider only visible range models (450-650nm) as this allows for non-contact imaging using standard cold surgical light sources \cite{Clancy2011}.
The visible light wavelength range also penetrates less deeply into tissue \cite{Eggert1987, Sabino2016}. Light-tissue interaction is therefore most likely to replicate a single homogeneous semi-infinite tissue layer and less likely to travel between layered structures.
%}
Tissue models tend to either be based on modifications of the Beer-Lambert model, or on the diffusion approximation particularly using the Kulbelka-Munk approximation method\cite{MacKenzie2018}. Three models that can be applied to homogeneous semi-infinite tissue in the visible wavelength range (450-650nm) include 1) a modified Beer-Lambert model \cite{Clancy2015}; 2) a more elaborate Beer-Lambert-based model proposed by Jacques \cite{Jacques1999}; and 3) a semi-empirical model based on the Kulbelka-Munk theory as described by Yudovsky \cite{Yudovsky2009}.
Alternatively, Monte Carlo (MC) methods may be used to model tissues for a range of wavelengths. Whilst MC is an established approach for modelling light-tissue interaction, it is computationally challenging to apply in an inverse problem setting making the estimation of optical properties from tissue spectra non trivial on the basis of MC simulations alone.
To address the need for improved inverse modelling utilising MC, Inverse Adding Doubling (IAD)\cite{Prahl2017} has been proposed, however this is also slow iterative process and therefore not as efficient as approaches based on analytical models.
A limitation of IAD is that, although a single reflectance measurement can be used to analyse semi-infinite samples with IAD, its primary focus is the extraction of optical properties from slabs of tissue with known thickness using both transmission and reflection measurements.
In this chapter, IAD is used to provide information on optical properties from tissue phantom slabs but is not assessed in the same capacity as the other 
%other purely
analytical models listed above.

Since there is no standard method of determining spatially resolved, ground truth, optical properties of tissues, 
simulations and physical
phantoms must be used
%to test analytical models.
for testing purposes.
Gelatin-based phantoms are commonly used to allow tunable absorption and scattering properties in solid phantoms \cite{Pogue2006, Gautam2023}.
%
In this chapter we compare three major analytical models against both MC simulated spectra, and measured spectra from gelatin-based tissue phantoms with known ground-truth constituent quantities. These datasets cover a large parameter range mimicking that expected in biological tissue\cite{Jacques2013}. Gelatin phantoms are constructed to optically mimic biological tissue absorption and scattering ranges as closely as possible, whilst maintaining well-defined optical properties with known ground truth. We present a first direct comparison in the performance of these models against simulated and measured spectra, and aim to evaluate their potential for accurate $StO_2$ extraction. This evaluation is quantified in terms of the forward models (utilising ground truth values within the models to predict the diffuse reflectance spectra), and for the inverse problem solving (evaluating the fidelity of the parameter extraction when fitting to the simulated or measured data).
Our results provide a comparison which can be used to inform the translation of these models into clinical applications.

\section{Materials and methods}\label{sec:methods}
\subsection{Tissue models}\label{sec:methodtissuemodels}
Three analytical models are compared and evaluated in this work: modified Beer-Lambert \cite{Clancy2015}, Jacques 1999 \cite{Jacques1999}, and Yudovsky 2009 \cite{Yudovsky2009}. Due to the absorbing and scattering nature of biological tissue these models are used to analyse diffuse reflectance spectra which result from propagation of light via many optical paths through the tissue.
% \begin{figure}
%     \centering
%     \includegraphics[width=0.4\textwidth]{TissueFigure.pdf}
%     \caption{Diagram depicting the multiple paths taken through tissue before the light exits to be detected as diffuse reflectance.}
%  \label{fig:tissuefigure}
% \end{figure}

\subsubsection{Common absorption and scattering model}\label{sec:opticproperties}
All three forward models use the wavelength dependent absorption and reduced scattering coefficients ($\mu_a(\lambda)$ and $\mu_s'(\lambda)$) of the tissue as inputs
%and output
to compute a
diffuse reflectance spectrum. 
The reduced scattering coefficient comprises of the scattering coefficient ($\mu_s(\lambda)$) and tissue anisotropy ($g$) as follows: 
\begin{equation}
    \mu_s'(\lambda) = \mu_s(\lambda) \times (1-g)
    \label{eq:reducedscattering}
\end{equation}
For biological tissue, the reduced scattering coefficient can be well approximated using Mie theory \cite{Jacques2013}: 
\begin{equation}
    \mu_s'(\lambda) = a(\frac{\lambda}{500})^{-b}
    \label{eq:Mie}
\end{equation}
where $a$ and $b$ are Mie scattering coefficients that range between 8\textrm{$cm^{-1}$} and 70\textrm{$cm^{-1}$}, and 0.1 and 3.3 respectively depending on the tissue microstructure \cite{Jacques2013}. 
The absorption coefficient of most internal, homogeneous, single-layer tissues is dominated by haemoglobin in the visible region (450-650 nm)\cite{JacquesAbs} and can be modelled using the following equation \cite{Yudovsky2009}: 
\begin{equation}
\begin{aligned}
    & \mu_a(\lambda) = f_{blood}\mu_{a, blood}(\lambda) + (1 - f_{blood})\mu_{a, back}(\lambda) \\
    & \textrm{where} \\
    & \mu_{a, blood}(\lambda) = c_{HbT}\frac{\ln(10)}{64500}[StO_2 \epsilon_{HbO_2}(\lambda) + (1 - StO_2)\epsilon_{Hb}(\lambda)] \\
    & \textrm{and} \\
    & \mu_{a, back}(\lambda) = 7.84\times10^8 \lambda^{-3.255}
\end{aligned}
\label{eq:mua}
\end{equation}
Here $StO_2$ refers to oxygen saturation, $c_{HbT}$ refers to the total concentration of haemoglobin in whole blood commonly taken as 150\textrm{$gL^{-1}$}\cite{Prahl1999}, with $StO_2$ denoting the fraction of this that is oxygenated ($HbO_2$) and the remainder is deoxygenated ($Hb$), and ranges between 0-100\%\cite{Yudovsky2009}. $f_{blood}$ is the volume fraction of tissue occupied by blood which has absorption coefficient $\mu_{a, blood}(\lambda)$ and is examined in the range of 0.2-7\%\cite{Yudovsky2009}, with the remainder of tissue having background absorption $\mu_{a, back}(\lambda)$\cite{Yudovsky2009}.
Finally, $\epsilon_{HbO_2}(\lambda)$ and $\epsilon_{Hb}(\lambda)$ denotes the wavelength-dependent extinction coefficients of the chromophores $HbO_2$ and $Hb$ which are found in the literature\cite{Prahl1999}.

\subsubsection{Modified Beer-Lambert}
The modified Beer-Lambert utilises these inputs to model absorption ($A(\lambda)$) and diffuse reflectance ($R(\lambda)$) as follows: 
\begin{equation}
\begin{aligned}
    & A(\lambda) = L\mu_a(\lambda) + \mu_s'(\lambda) \\
    & R(\lambda) = \exp{\left(-\frac{A(\lambda)}{100}\right)}
\end{aligned}
\label{eq:modBL1}
\end{equation}
Conventionally, $\mu_a(\lambda)$ and $\mu_s'(\lambda)$ are quoted in units of cm\textrm{$^{-1}$}, however in this model units of mm\textrm{$^{-1}$} are required for diffuse reflectance units of \% hence a factor of 100 is included. $L$ describes a differential path length to account for the variety of photon path lengths through scattering media. $L$ is often simplified to be equal to 1 \cite{Clancy2015} and $\mu_s'(\lambda)$ is often modelled as a wavelength-independent constant\cite{Clancy2015, Ma2016}. 
To allow for further flexibility in this model,
we introduce linear scaling hyperparameters ($M_{1-3}$), as shown in Equation \eqref{eq:modBL2}, that can be fitted to MC simulations at the refractive index of interest. 
\begin{equation}
\begin{aligned}
    & A(\lambda) = M_1\mu_a(\lambda) + M_2\mu_s'(\lambda) + M_3 \\
    & R(\lambda) = \exp{\left(-\frac{A(\lambda)}{100}\right)}
\end{aligned}
\label{eq:modBL2}
\end{equation}

\subsubsection{Jacques 1999}\label{sec:Jacques}
The Jacques model is also based on the Beer-Lambert model, where an assumption is made that the ensemble of path lengths experienced by photons in a tissue can be approximated by a single wavelength-dependent path length $L(\lambda) = A(\lambda)\delta(\lambda)$ where $A(\lambda)$ and $\delta(\lambda)$ are defined in Equation \ref{eq:Jacques}. This results in the following model with the hyperparameters ($M_{1-3}$) which the authors
%state are fitted
fit
to Adding Doubling simulations \cite{Jacques1999}. In our work, we refit these to MC simulations since these are considered the gold standard optical simulation method and improve the model fitting results. 
\begin{equation}
\begin{aligned}
    & N'(\lambda) = \frac{\mu_s'(\lambda)}{\mu_a(\lambda)} \\
    & \delta(\lambda) = \frac{1}{\sqrt{3\mu_a(\lambda)[\mu_a(\lambda) + \mu_s'(\lambda)]}} \\
    & A(\lambda) = M_1 + M_2\exp \left[ \frac{\ln(N'(\lambda))}{M_3} \right] \\
    & R(\lambda) = \exp[-A(\lambda)\delta(\lambda)\mu_a(\lambda)] \\
\end{aligned}
\label{eq:Jacques}
\end{equation}

\subsubsection{Yudovsky 2009}\label{sec:Yudovskysingle}
The original Yudovsky model presents a complex analytical derivation \cite{Yudovsky2009} with a simplified formulation subsequently presented in their Erratum \cite{Yudovsky2015}. The model takes the reduced albedo ($w'(\lambda)$) as input and returns the diffuse reflectance spectra ($R(\lambda)$). This provides an easily applicable model with hyperparameters ($M_{1-6}$) which are quoted for a refractive index ($n$) of 1.44. We found that these can be fitted to MC spectra to allow for use with other refractive indices. 
\begin{equation}
\begin{aligned}
    & w'(\lambda) = \frac{\mu_s'(\lambda)}{\mu_a(\lambda) + \mu_s'(\lambda)} \\
    & R = M_1 + M_2\exp{\left[ M_3w'(\lambda)^{M_4}\right]} + \frac{M_5}{1.02 - M_6} \\
\end{aligned}
\label{eq:Yudovskysingle}
\end{equation}

@@

\subsubsection{Monte Carlo simulations}\label{sec:methodsMC}
The models take $\mu_a(\lambda)$ and $\mu_s'(\lambda)$ as inputs. These are given for bulk tissue in Equations \eqref{eq:mua} and \eqref{eq:Mie} and used to generate MC simulated reference spectra. MC takes $\mu_s(\lambda)$ as input, not $\mu_s'(\lambda)$, so a random value of anisotropy ($g$) is chosen per spectrum between 0.7-0.9\cite{Yudovsky2009} and used in Equation \eqref{eq:reducedscattering} for conversion to $\mu_s(\lambda)$. The variable parameters are $a$, $b$, $StO_2$, and $f_{blood}$ which are bounded to 8-70cm\textrm{$^{-1}$}, 0.1-3.3, 0-100\%, 0.2-7\% respectively to mimic biological tissue\cite{Yudovsky2009, Jacques2013}; a value of each of these variables is randomly selected within these bounds to generate each of the 100 simulated spectra, meaning 100 values of each parameter are sampled. When fitting the inverse models to these spectra for parameter extraction (as described in Section \ref{sec:methodevaluate}), the same bounds are imposed on the fitting routine. These simulated spectra are generated using CUDAMCML\cite{Alerstam2008}, which is a GPU accelerated adaptation of the well-established MCML programme\cite{Wang1995}. It has been shown that the semi-infinite approximation is valid for thicknesses of above 1cm\cite{Zhang2014} so 100 spectra were simulated for each refractive index of 1.33, 1.35, and 1.44, by propagating 100,000 photons through the semi-infinite slabs approximated by a thickness of 3cm. The refractive indices were chosen to represent the common phantom and tissue use-cases of these models: 1.33 for use with water-based phantoms, 1.35 for use with gelatin-based phantoms (as in this work), and 1.44 for use with biological tissue.

@@

The dyes acid red 1 (AR1, 210633, Merck, Germany) and acid red 14 (AR14, B22328, Fisher Scientific, England) are chosen to mimic the extinction coefficients of oxygenated and deoxygenated haemoglobin respectively, with a third dye of crystal violet (CV, C6158, Merck, Germany) chosen to investigate the effect of including further chromophores. As in the modelled tissue $\mu_a(\lambda)$ (Equation \eqref{eq:mua}), the total dye concentration is modelled independently of the relative ratios of each dye. To ensure the absorbance impact of each dye is approximately equal, a factor of $\frac{5}{3}$ is included for AR14 and $\frac{1}{2}$ for CV computationally by combining with the extinction coefficients to create effective extinction coefficients ($\epsilon_{eff}$) that have approximately equal impact. To reflect this experimentally, the concentrations of these dyes are modified by these factors. These effective extinction coefficients calculated from literature values\cite{PhotochemCAD}, alongside those of haemoglobin\cite{Prahl1998}, can be seen in Figure \ref{fig:phantommethods}a. 
% \begin{figure}[h]
%     \centering
%     \includegraphics[width=0.6\textwidth]{effective_eps_lit.pdf}
%     \caption{Effective extinction coefficients of acid red 1 (AR1), acid red 14 (AR14), and crystal violet (CV) displayed with the extinction coefficients of oxygenated (HbO$_2$) and deoxygenated haemoglobin (Hb).}
%  \label{fig:exteff}
% \end{figure}

@@

Intralipid is used to modulate the scattering coefficient of these phantoms which can be modelled with a Mie scattering function (Equation \eqref{eq:Mie}) using parameters within the range of tissues. 5 volume fractions of intralipid are chosen between 1\% and 6\% to cover a range of scattering parameters within those seen in biological tissue \cite{Jacques2013}.

Each phantom consists of 6\% gelatin (Type A approximately 175g bloom, G2625, Merck, Germany) by mass %to mimic the properties of healthy tissue \cite{Doyley2003} 
and 0.5\% of 4\% formaldehyde (J60401, Fisher Scientific, England) to increase the melting point of the phantoms and allow their use at room temperature \cite{Pogue2006}. The remainder of each phantom consists of the dye solutions in a variety of ratios combined with intralipid at a variety of concentrations.

@@

Moulds with a depth of 1 cm are used to allow measurement of diffuse reflectance within a semi-infinite regime\cite{Zhang2014}.
MC simulations were used to confirm that this thickness was sufficient to be within the semi-infinite regime.
Some additional phantoms are constructed with a depth of 5 mm to allow for sufficient transmission for total transmittance measurements required for inverse adding doubling (IAD) analysis \cite{Prahl2017}. 
%to extract $\mu_a$ and $\mu_s'$ from total reflectance and diffuse transmittance measurements.

>>>>

\begin{center}
\begin{minipage}[b]{0.9\linewidth}
\small
\textbf{Foreword\,}
This chapter is an \emph{in extenso} reproduction of~\cite{Bahl2023a} and has been submitted to Journal of Biophotonics for publication. 
\newline
AB collaborated with TV to identify the analytical tissue models for comparison. AB 
%identified the three major analytical tissue models for comparison, 
designed and conducted the simulations and synthetic experiments and measurements to construct the datasets for evaluation, and conducted the evaluation of the models under the supervision of JS, MB, and TV. AB worked collaboratively with SS and YX to identify and validate the appropriate IAD method to use in this work. 
\end{minipage}
\end{center}

@@

% \maketitle 
% 
\section{Introduction}\label{sec:intro}
As mentioned in Chapter \ref{chap:intro}, $StO_2$ 
%Tissue oxygen saturation ($StO_2$) 
is a key metric that could be used in surgery to determine tissue viability, tumour localisation, and vessel flow~\cite{Takami2017, Hughes2019, Richardson2023}.
There are, however, no currently established, spatially-resolved, intra-operative methods of quantitatively determining $StO_2$ and so perfusion is often displayed (e.g. using DSA, MVD, or ICG) or global $StO_2$ is discussed from JVOS or cerebral oximetry.
% In many clinical applications indocyanine green fluorescence angiography (ICG) is used to display perfusion as 
% %oxygenated
% blood flow through vessels can be visualised by the fluorescence of the indocyanine dye given to the patient \cite{Hackethal2018}.
% Whilst this method has shown clinical benefits \cite{Renna2023, Teng2021}, it cannot quantify the $StO_2$ of the tissues being visualised leading to subjective interpretation.
There has been increasing promise that diffuse reflectance spectroscopic techniques could be used to obtain $StO_2$ intra-operatively. Hyperspectral Imaging (HSI) is one such technique that obtains a diffuse reflectance spectrum for each pixel of an image~\cite{Kulcke2018, Taylor-Williams2022}, enabling intra-operative, non-contact, spatially-resolved $StO_2$ extraction without the need for a contrast agent.

There have been many proposed methods to extract $StO_2$ parameters from tissue diffuse reflectance spectra ranging in complexity, as mentioned in Chapter \ref{chap:intro}.
%from the simplest ratiometric two-wavelength methods to more sophisticated analytical models \cite{MacKenzie2018} or deep-learning based approaches \cite{Ayala2023}. Using two or three wavelength models can provide good results but require data to be captured at these precise wavelengths placing large constraints on the measurement devices used and significant model assumptions. 
Analytical models can be applied to ranges of wavelengths for which they are developed and so these are favoured in this work. 
%\revref{2}{6}\textcolor{red}{
Here we consider only visible range models (450-650nm) as this allows for non-contact imaging using standard cold surgical light sources~\cite{Clancy2011}.
The visible light wavelength range also penetrates less deeply into tissue~\cite{Eggert1987, Sabino2016}. Light-tissue interaction is therefore most likely to replicate a single homogeneous semi-infinite tissue layer and less likely to travel between layered structures.
%}
Tissue models tend to either be based on modifications of the Beer-Lambert model, or on the diffusion approximation particularly using the Kulbelka-Munk approximation method~\cite{MacKenzie2018}. Three models that can be applied to homogeneous semi-infinite tissue in the visible wavelength range (450-650nm) include 1) a modified Beer-Lambert model~\cite{Clancy2015}; 2) a more elaborate Beer-Lambert-based model proposed by Jacques~\cite{Jacques1999}; and 3) a semi-empirical model based on the Kulbelka-Munk theory as described by Yudovsky~\cite{Yudovsky2009}.
Alternatively, Monte Carlo (MC) methods may be used to model tissues for a range of wavelengths. Whilst MC is an established approach for modelling light-tissue interaction, it is computationally challenging to apply in an inverse problem setting making the estimation of optical properties from tissue spectra non trivial on the basis of MC simulations alone.
To address the need for improved inverse modelling utilising MC, Inverse Adding Doubling (IAD)~\cite{Prahl2017} has been proposed, however this is also slow iterative process and therefore not as efficient as approaches based on analytical models.
A limitation of IAD is that, although a single reflectance measurement can be used to analyse semi-infinite samples with IAD, its primary focus is the extraction of optical properties from slabs of tissue with known thickness using both transmission and reflection measurements.
In this chapter, IAD is used to provide information on optical properties from tissue phantom slabs but is not assessed in the same capacity as the other 
%other purely
analytical models listed above.

Since there is no standard method of determining spatially resolved, ground truth, optical properties of tissues, 
simulations and physical
phantoms must be used
%to test analytical models.
for testing purposes.
Gelatin-based phantoms are commonly used to allow tunable absorption and scattering properties in solid phantoms~\cite{Pogue2006, Gautam2023}.
%
In this chapter we compare three major analytical models against both MC simulated spectra, and measured spectra from gelatin-based tissue phantoms with known ground-truth constituent quantities. These datasets cover a large parameter range mimicking that expected in biological tissue~\cite{Jacques2013}. Gelatin phantoms are constructed to optically mimic biological tissue absorption and scattering ranges as closely as possible, whilst maintaining well-defined optical properties with known ground truth. We present a first direct comparison in the performance of these models against simulated and measured spectra, and aim to evaluate their potential for accurate $StO_2$ extraction. This evaluation is quantified in terms of the forward models (utilising ground truth values within the models to predict the diffuse reflectance spectra), and for the inverse problem solving (evaluating the fidelity of the parameter extraction when fitting to the simulated or measured data).
Our results provide a comparison which can be used to inform the translation of these models into clinical applications.

\section{Materials and methods}\label{sec:methods}
\subsection{Tissue models}\label{sec:methodtissuemodels}
Three analytical models are compared and evaluated in this work: modified Beer-Lambert~\cite{Clancy2015}, Jacques 1999~\cite{Jacques1999}, and Yudovsky 2009~\cite{Yudovsky2009}. Due to the absorbing and scattering nature of biological tissue these models are used to analyse diffuse reflectance spectra which result from propagation of light via many optical paths through the tissue.
% \begin{figure}
%     \centering
%     \includegraphics[width=0.4\textwidth]{TissueFigure.pdf}
%     \caption{Diagram depicting the multiple paths taken through tissue before the light exits to be detected as diffuse reflectance.}
%  \label{fig:tissuefigure}
% \end{figure}

\subsubsection{Common absorption and scattering model}\label{sec:opticproperties}
All three forward models use the wavelength dependent absorption and reduced scattering coefficients ($\mu_a(\lambda)$ and $\mu_s'(\lambda)$) of the tissue as inputs
%and output
to compute a
diffuse reflectance spectrum. 
The reduced scattering coefficient comprises of the scattering coefficient ($\mu_s(\lambda)$) and tissue anisotropy ($g$) as follows: 
\begin{equation}
    \mu_s'(\lambda) = \mu_s(\lambda) \times (1-g)
    \label{eq:reducedscattering}
\end{equation}
For biological tissue, the reduced scattering coefficient can be well approximated using Mie theory~\cite{Jacques2013}: 
\begin{equation}
    \mu_s'(\lambda) = a(\frac{\lambda}{500})^{-b}
    \label{eq:Mie}
\end{equation}
where $a$ and $b$ are Mie scattering coefficients that range between 8\textrm{$cm^{-1}$} and 70\textrm{$cm^{-1}$}, and 0.1 and 3.3 respectively depending on the tissue microstructure~\cite{Jacques2013}. 
The absorption coefficient of most internal, homogeneous, single-layer tissues is dominated by haemoglobin in the visible region (450-650 nm)~\cite{JacquesAbs} and can be modelled using the following equation~\cite{Yudovsky2009}: 
\begin{equation}
\begin{aligned}
    & \mu_a(\lambda) = f_{blood}\mu_{a, blood}(\lambda) + (1 - f_{blood})\mu_{a, back}(\lambda) \\
    & \textrm{where} \\
    & \mu_{a, blood}(\lambda) = c_{HbT}\frac{\ln(10)}{64500}[StO_2 \epsilon_{HbO_2}(\lambda) + (1 - StO_2)\epsilon_{Hb}(\lambda)] \\
    & \textrm{and} \\
    & \mu_{a, back}(\lambda) = 7.84\times10^8 \lambda^{-3.255}
\end{aligned}
\label{eq:mua}
\end{equation}
Here $StO_2$ refers to oxygen saturation, $c_{HbT}$ refers to the total concentration of haemoglobin in whole blood commonly taken as 150\textrm{$gL^{-1}$}~\cite{Prahl1999}, with $StO_2$ denoting the fraction of this that is oxygenated ($HbO_2$) and the remainder is deoxygenated ($Hb$), and ranges between 0-100\%~\cite{Yudovsky2009}. $f_{blood}$ is the volume fraction of tissue occupied by blood which has absorption coefficient $\mu_{a, blood}(\lambda)$ and is examined in the range of 0.2-7\%~\cite{Yudovsky2009}, with the remainder of tissue having background absorption $\mu_{a, back}(\lambda)$~\cite{Yudovsky2009}.
Finally, $\epsilon_{HbO_2}(\lambda)$ and $\epsilon_{Hb}(\lambda)$ denotes the wavelength-dependent extinction coefficients of the chromophores $HbO_2$ and $Hb$ which are found in the literature~\cite{Prahl1999}.

\subsubsection{Modified Beer-Lambert}
The modified Beer-Lambert utilises these inputs to model absorption ($A(\lambda)$) and diffuse reflectance ($R(\lambda)$) as follows: 
\begin{equation}
\begin{aligned}
    & A(\lambda) = L\mu_a(\lambda) + \mu_s'(\lambda) \\
    & R(\lambda) = \exp{\left(-\frac{A(\lambda)}{100}\right)}
\end{aligned}
\label{eq:modBL1}
\end{equation}
Conventionally, $\mu_a(\lambda)$ and $\mu_s'(\lambda)$ are quoted in units of cm\textrm{$^{-1}$}, however in this model units of mm\textrm{$^{-1}$} are required for diffuse reflectance units of \% hence a factor of 100 is included. $L$ describes a differential path length to account for the variety of photon path lengths through scattering media. $L$ is often simplified to be equal to 1~\cite{Clancy2015} and $\mu_s'(\lambda)$ is often modelled as a wavelength-independent constant~\cite{Clancy2015, Ma2016}. 
To allow for further flexibility in this model,
we introduce linear scaling hyperparameters ($M_{1-3}$), as shown in Equation \eqref{eq:modBL2}, that can be fitted to MC simulations at the refractive index of interest. 
\begin{equation}
\begin{aligned}
    & A(\lambda) = M_1\mu_a(\lambda) + M_2\mu_s'(\lambda) + M_3 \\
    & R(\lambda) = \exp{\left(-\frac{A(\lambda)}{100}\right)}
\end{aligned}
\label{eq:modBL2}
\end{equation}

\subsubsection{Jacques 1999}\label{sec:Jacques}
The Jacques model is also based on the Beer-Lambert model, where an assumption is made that the ensemble of path lengths experienced by photons in a tissue can be approximated by a single wavelength-dependent path length $L(\lambda) = A(\lambda)\delta(\lambda)$ where $A(\lambda)$ and $\delta(\lambda)$ are defined in Equation \ref{eq:Jacques}. This results in the following model with the hyperparameters ($M_{1-3}$) which the authors
%state are fitted
fit
to Adding Doubling simulations~\cite{Jacques1999}. In our work, we refit these to MC simulations since these are considered the gold standard optical simulation method and improve the model fitting results. 
\begin{equation}
\begin{aligned}
    & N'(\lambda) = \frac{\mu_s'(\lambda)}{\mu_a(\lambda)} \\
    & \delta(\lambda) = \frac{1}{\sqrt{3\mu_a(\lambda)[\mu_a(\lambda) + \mu_s'(\lambda)]}} \\
    & A(\lambda) = M_1 + M_2\exp \left[ \frac{\ln(N'(\lambda))}{M_3} \right] \\
    & R(\lambda) = \exp[-A(\lambda)\delta(\lambda)\mu_a(\lambda)] \\
\end{aligned}
\label{eq:Jacques}
\end{equation}

\subsubsection{Yudovsky 2009}\label{sec:Yudovskysingle}
The original Yudovsky model presents a complex analytical derivation~\cite{Yudovsky2009} with a simplified formulation subsequently presented in their Erratum~\cite{Yudovsky2015}. The model takes the reduced albedo ($w'(\lambda)$) as input and returns the diffuse reflectance spectra ($R(\lambda)$). This provides an easily applicable model with hyperparameters ($M_{1-6}$) which are quoted for a refractive index ($n$) of 1.44. We found that these can be fitted to MC spectra to allow for use with other refractive indices. 
\begin{equation}
\begin{aligned}
    & w'(\lambda) = \frac{\mu_s'(\lambda)}{\mu_a(\lambda) + \mu_s'(\lambda)} \\
    & R = M_1 + M_2\exp{\left[ M_3w'(\lambda)^{M_4}\right]} + \frac{M_5}{1.02 - M_6} \\
\end{aligned}
\label{eq:Yudovskysingle}
\end{equation}

@@

\subsubsection{Monte Carlo simulations}\label{sec:methodsMC}
The models take $\mu_a(\lambda)$ and $\mu_s'(\lambda)$ as inputs. These are given for bulk tissue in Equations \eqref{eq:mua} and \eqref{eq:Mie} and used to generate MC simulated reference spectra. MC takes $\mu_s(\lambda)$ as input, not $\mu_s'(\lambda)$, so a random value of anisotropy ($g$) is chosen per spectrum between 0.7-0.9~\cite{Yudovsky2009} and used in Equation \eqref{eq:reducedscattering} for conversion to $\mu_s(\lambda)$. The variable parameters are $a$, $b$, $StO_2$, and $f_{blood}$ which are bounded to 8-70cm\textrm{$^{-1}$}, 0.1-3.3, 0-100\%, 0.2-7\% respectively to mimic biological tissue~\cite{Yudovsky2009, Jacques2013}; a value of each of these variables is randomly selected within these bounds to generate each of the 100 simulated spectra, meaning 100 values of each parameter are sampled. When fitting the inverse models to these spectra for parameter extraction (as described in Section \ref{sec:methodevaluate}), the same bounds are imposed on the fitting routine. These simulated spectra are generated using CUDAMCML~\cite{Alerstam2008}, which is a GPU accelerated adaptation of the well-established MCML programme~\cite{Wang1995}. It has been shown that the semi-infinite approximation is valid for thicknesses of above 1cm~\cite{Zhang2014} so 100 spectra were simulated for each refractive index of 1.33, 1.35, and 1.44, by propagating 100,000 photons through the semi-infinite slabs approximated by a thickness of 3cm. The refractive indices were chosen to represent the common phantom and tissue use-cases of these models: 1.33 for use with water-based phantoms, 1.35 for use with gelatin-based phantoms (as in this work), and 1.44 for use with biological tissue.

@@

The dyes acid red 1 (AR1, 210633, Merck, Germany) and acid red 14 (AR14, B22328, Fisher Scientific, England) are chosen to mimic the extinction coefficients of oxygenated and deoxygenated haemoglobin respectively, with a third dye of crystal violet (CV, C6158, Merck, Germany) chosen to investigate the effect of including further chromophores. As in the modelled tissue $\mu_a(\lambda)$ (Equation \eqref{eq:mua}), the total dye concentration is modelled independently of the relative ratios of each dye. To ensure the absorbance impact of each dye is approximately equal, a factor of $\frac{5}{3}$ is included for AR14 and $\frac{1}{2}$ for CV computationally by combining with the extinction coefficients to create effective extinction coefficients ($\epsilon_{eff}$) that have approximately equal impact. To reflect this experimentally, the concentrations of these dyes are modified by these factors. These effective extinction coefficients calculated from literature values~\cite{PhotochemCAD}, alongside those of haemoglobin~\cite{Prahl1998}, can be seen in Figure \ref{fig:phantommethods}a. 
% \begin{figure}[h]
%     \centering
%     \includegraphics[width=0.6\textwidth]{effective_eps_lit.pdf}
%     \caption{Effective extinction coefficients of acid red 1 (AR1), acid red 14 (AR14), and crystal violet (CV) displayed with the extinction coefficients of oxygenated (HbO$_2$) and deoxygenated haemoglobin (Hb).}
%  \label{fig:exteff}
% \end{figure}

@@

Intralipid is used to modulate the scattering coefficient of these phantoms which can be modelled with a Mie scattering function (Equation \eqref{eq:Mie}) using parameters within the range of tissues. 5 volume fractions of intralipid are chosen between 1\% and 6\% to cover a range of scattering parameters within those seen in biological tissue~\cite{Jacques2013}.

Each phantom consists of 6\% gelatin (Type A approximately 175g bloom, G2625, Merck, Germany) by mass %to mimic the properties of healthy tissue \cite{Doyley2003} 
and 0.5\% of 4\% formaldehyde (J60401, Fisher Scientific, England) to increase the melting point of the phantoms and allow their use at room temperature~\cite{Pogue2006}. The remainder of each phantom consists of the dye solutions in a variety of ratios combined with intralipid at a variety of concentrations.

@@

Moulds with a depth of 1 cm are used to allow measurement of diffuse reflectance within a semi-infinite regime~\cite{Zhang2014}.
MC simulations were used to confirm that this thickness was sufficient to be within the semi-infinite regime.
Some additional phantoms are constructed with a depth of 5 mm to allow for sufficient transmission for total transmittance measurements required for inverse adding doubling (IAD) analysis~\cite{Prahl2017}. 
%to extract $\mu_a$ and $\mu_s'$ from total reflectance and diffuse transmittance measurements.
